Safe and Robust Formal Synthesis of Digital Controllers for Continuous Plants with Transient Performance Specifications

### 2017-05-24 automatica/automatica.tex
added relevant appendix
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

%\appendix
%\section{A summary of Latin grammar}    % Each appendix must have a short title.
%\section{Some Latin vocabulary}         % Sections and subsections are supported  
                                        % in the appendices.
\end{document}

>>>>

\appendix

%-------------------------------
\section{Stability of Closed-loop Models}
\label{sec:appendix-stability}
%-------------------------------

%-------------------------------
\subsection{Stability of closed-loop models with fixed-point controller error}
\label{sec:stab_FWL}
%-------------------------------

The proof of Jury's criterion~\cite{fadali} relies on the fact that the relationship between states and
next states is defined by $x_{k+1} = (A_d - B_dK) x_k$, all computed
at infinite precision.  When we employ a FWL digital controller, the
operation becomes:
%
\begin{align*}
x_{k+1} &= A_d \cdot x_{k} -(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})).  \\
x_{k+1} &= (A_d  - B_dK) \cdot x_k + B_dK\delta, 
\end{align*}
%
where $\delta$ is the maximum error that can be introduced by the FWL
controller in one step, i.e., by reading the states values once and
multiplying by $K$ once.  We derive the closed form expression for $x_n$ as
follows:
%
\begin{align*}
x_{1} &= (A_d  - B_dK)x_0 + B_dK\delta \\
x_{2} 
 &=(A_d  - B_dK)^2x_0 + (A_d  - B_dK)B_dK\delta + B_dK\delta \\
x_{n} &= (A_d  - B_dK)^nx_0 + (A_d  - B_dK)^{n-1}B_dK\delta + ... \\  \nonumber & + (A_d  - B_dK)^1B_dK \delta + B_dK\delta \\
  &= (A_d - B_dK)^nx_0 + \sum_{i=0}^{i=n-1}(A_d - B_dK)^iB_dk\delta. 
\end{align*}

The definition of asymptotic stability is that the system converges to a
reference signal, in this case we use no reference signal so an
asymptotically stable system will converge to the origin.  We know that the
original system with an infinite-precision controller is stable, because we
have synthesized it to meet Jury's criterion.  Hence, $(A_d - B_dK)^n x_0$ 
must converge to zero.

The power series of matrices
converges~\cite{horn1990matrix} iff the eigenvalues of the matrix are less
than~1 as follows:
%
%\begin{align*}
$\sum_{i=0}^{\infty}T^i  = (I - T)^{-1}$, 
%\end{align*}
%
where $I$ is the identity matrix and $T$ is a square matrix. Thus, our system will converge to the value 
%
\begin{align*}
0 + (I - A_d + B_dK)^{-1}B_dk\delta \,. 
\end{align*}
%
As a result, if the value $(I - A_d + B_dK)^{-1}B_dk\delta$ is within the
safe space, then the synthesized fixed-point controller results in a safe
closed-loop model.  The convergence to a finite value, however, will not
make it asymptomatically stable.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Errors in LTI models} \label{sec:appendix:LTIbackground}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

\subsection{Errors due to numerical representation} \label{appendix:numerical_errors}

We have  used $\mathcal{F}_{\langle I,F \rangle}(x)$ denote a real number
$x$ represented in a fixed point domain, with $I$ bits representing the
integer part and $F$ bits representing the decimal part.  The smallest
number that can be represented in this domain is $c_m=2^{-F}$.  The
following approximation errors will arise in mathematical operations and
representation:
%
\begin{enumerate}

\item {\bf Truncation:} Let $x$ be a real number, and $\mathcal{F}_{\langle
I,F \rangle}(x)$ be the same number represented in a fixed-point domain as
above.  Then $\mathcal{F}_{\langle I,F \rangle}(x) = x-\delta_T$ where the
error $ \delta_T=x\ \%_{c_m}\ \tilde x$, and $\%_{c_m}$ is the modulus
operation performed on the last bit. 
Thus, $\delta_T$ is the truncation error and it will propagate across
operations.
%
\item {\bf Rounding:} The following errors appear in basic operations.  Let
$c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be
the truncation errors caused by representing $c_1$ and $c_2$ in the
fixed-point domain as above.
%
\begin{enumerate}
%
\item Addition/Subtraction: these operations only propagate errors coming
from truncation of the operands, namely $\mathcal{F}_{\langle I,F
\rangle}(c_1) \pm \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_3$
with $|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.
%
\item Multiplication: $\mathcal{F}_{\langle I,F \rangle}(c_1) \cdot
\mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3$ with $|\delta_3|
\leq |\delta_{T1}\cdot\mathcal{F}_{\langle I,F \rangle}(c_2)|\allowbreak +
|\delta_{T2}\cdot\mathcal{F}_{\langle I,F \rangle}(c_1)| + c_m$, where
$c_m=2^{-F}$ as above.
%
\item Division: the operations performed by our controllers in the FWL
domain do not include division.  However, we do use division in computations
at the precision of the plant.  Here the error depends on whether the
divisor is greater or smaller than the dividend:  $\mathcal{F}_{\langle I,F
\rangle}(c_1) / \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_{T3}$
where $\delta_{T3}$ is $(\delta_{T2}\cdot c_1 - \delta_{T1}\cdot
c_2)/(\delta_{T2}^2 - \delta_{T2} c_2)$,
\end{enumerate}

\item {\bf Overflow:}
The maximum size of a real number $x$ that can be represented in a fixed
point domain as $\mathcal{F}_{\langle I,F \rangle}(x)$ is $\pm
(2^{I-1}+1-2^{-F})$.  Numbers outside this range cannot be represented by
the domain.  We check that overflow does not occur.

\end{enumerate}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Modeling quantization as noise} \label{appendix:quantization-noise}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

During any given ADC conversion, the continuous signal will be sampled in
the real domain and transformed by $\mathcal{F}\langle I_{c},F_{c} \rangle
(x)$ (assuming the ADC discretization is the same as the digital
implementation).  This sampling uses a threshold which is defined by the
less significant bit ($q_{c}=c_{m_c}=2^{-F_c}$) 
of the ADC and some non-linearities of the circuitry.  The overall conversion is
%
$$\mathcal{F}\langle I_{c},F_{c} \rangle(y(t)) = y_k :
y_k \in \left[y(t)-\frac{q_{c}}{2}\ \ \ \ y(t)+\frac{q_{c}}{2}\right] \,.$$
%
If we denote the error in the conversion by $\nu_k=y_k-y(t)$ where $t = nk$,
and $n$ is the sampling time and $k$ the number of steps, then we may define
some bounds for it $\nu_k \in [-\frac{q_{c}}{2}\ \ \frac{q_{c}}{2}]$.

We will assume, for the purposes of this analysis, that the domain of the
ADC is that of the digital controller (i.e, the quantizer includes any
digital gain added in the code).  The process of quantization in the DAC is
similar except that it is calculating $\mathcal{F}\langle I_{dac},F_{dac}
\rangle (\mathcal{F}\langle I_{c},F_{c} \rangle (x)) $.  If these domains
are the same ($I_{c}=I_{dac},\allowbreak F_{c}=F_{dac}$), or if the DAC
resolution in higher than the ADCs, then the DAC quantization error is equal
to zero.  From the above equations we can now define the ADC and DAC
quantization noises ${\nu_1}_k \in [-\frac{q_1}{2}\ \ \frac{q_1}{2}]$ and
${\nu_2}_k \in [-\frac{q_2}{2}\ \ \frac{q_2}{2}]$, where $q_1=q_{c}$ and 
$q_2=q_\mathit{dac}$.  This is illustrated in
Fig.~\ref{ssec:statefeedbackcontrol} where $Q_1$ is the quantizer of the ADC
and $Q_2$ the quantizer for the DAC.  These bounds hold irrespective of
whether the noise is correlated, hence we may use them to over-approximate
the noise effect on the state space progression over time.  The
resulting dynamics are
%
\begin{align*}
%\label{eq:pre_quantization}
{x}_{k+1} = {A}_d{x}_k+{B}_d({u}_k+{{\nu}_2}_k), \quad u_k = -K{x}_{k}+{{\nu}_1}_k, 
\end{align*}
%
which result in the following closed-loop dynamics:
%
\begin{align*}
%\label{eq:quantization}
{x}_{k+1} &= ({A}_d-{B}_d{K}_d) {x}_k+{B}_d{{\nu}_2}_k +{{\nu}_1}_k \,. 
\end{align*}

%\section{A summary of Latin grammar}    % Each appendix must have a short title.
%\section{Some Latin vocabulary}         % Sections and subsections are supported  
                                        % in the appendices.
\end{document}
